### 2016-10-06 paper/7_conclusion.tex
..

Gilbert tackles the problem of scalable analytics by fusing the assets of high-level linear algebra abstractions with the power of massively parallel dataflow systems. Gilbert is a fully functional linear algebra environment, which is programmed by the widespread MATLAB language. Gilbert programs are executed on massively parallel dataflow systems, which allow to process data exceeding the capacity of a single computer's memory. The system itself comprises the technology stack to parse, type and compile MATLAB code into a format which can be executed in parallel. In order to apply high-level linear algebra optimizations, we conceived an intermediate representation for linear algebra programs. Gilbert's optimizer exploits this representation to remove redundant transpose operations and to determine an optimal matrix multiplication order. The optimized program is translated into an highly optimized execution plan suitable for the execution on a supported engine. Gilbert allows the distributed execution on Spark and Flink. Our systematical evaluation has shown that Gilbert supports all fundamental linear algebra operations, making it fully operational.
Additionally, its loop support allows to implement a wide multitude of machine learning algorithms.
Exemplary, we have implemented the PageRank and GNMF algorithm.
The code changes required to make these algorithms run in Gilbert are minimal and only concern Gilbert's loop abstraction.
Our benchmarks have proven that Gilbert can handle data sizes which no longer can be efficiently processed on a single computer.
Moreover, Gilbert showed a promising scale out behavior, making it a suitable candidate for large-scale data processing.

>>>>

Gilbert addresses the problem of scalable analytics by fusing the assets of high-level linear algebra abstractions with the power of massively parallel dataflow systems. Gilbert is a fully functional linear algebra environment, which is programmed by the widespread MATLAB language. Gilbert programs are executed on massively parallel dataflow systems, which allow to process data exceeding the capacity of a single computer's memory. The system itself comprises the technology stack to parse, type and compile MATLAB code into a format which can be executed in parallel. In order to apply high-level linear algebra optimizations, we conceived an intermediate representation for linear algebra programs. Gilbert's optimizer exploits this representation to remove redundant transpose operations and to determine an optimal matrix multiplication order. The optimized program is translated into an highly optimized execution plan suitable for the execution on a supported engine. Gilbert allows the distributed execution on Spark and Flink. Our systematical evaluation has shown that Gilbert supports all fundamental linear algebra operations, making it fully operational.
Additionally, its loop support allows to implement a wide multitude of machine learning algorithms.
Exemplary, we have implemented the PageRank and GNMF algorithm.
The code changes required to make these algorithms run in Gilbert are minimal and only concern Gilbert's loop abstraction.
Our benchmarks have proven that Gilbert can handle data sizes which no longer can be efficiently processed on a single computer.
Moreover, Gilbert showed a promising scale out behavior, making it a suitable candidate for large-scale data processing.
